\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data

### 2016-07-18 article.tex
kinship2 format consistency
title case and caps after colon

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
    \caption{Example pedigree chart from the kinship2 package, where the vertical axis denotes generation count. We superimposed green-highlighted individual 215 for explanatory purposes. As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual. Hence, it should be displayed twice on the vertical axis, once for each of its generation counts. However, standard pedigree tools only allow for an individual to be displayed once, leading to ambiguous portrayal of the genealogical dataset.}
    \label{fig:kinshipFig}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{pathTNZB}}
    \caption{Left: The shortest path between varieties Tokyo and Narow is strictly composed of a unidirectional sequence of parent-child relationships. Right: The shortest path between varieties Zane and Bedford is not strictly composed of unidirectional parent-child relationships; they instead have a cousin-like relationship.}
    \label{fig:pathTNZB}
\end{figure}

@@

\begin{figure}%[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotTNBin3}}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
    \label{fig:plotTNBin3}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotTNBin6}}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
    \label{fig:plotTNBin6}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{LeeAD}}
    \caption{Top: All ancestors and descendants of the variety Lee are shown in this \pkg{ggenealogy} plot. Bottom: We now attempt to mimic the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated. The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee. The presence of red parentheses indicate that the plot on the bottom ambiguously display the example soybean genealogy in the way that the \pkg{ggenealogy} plot on the top can accomplish.}
    \label{fig:Lee}
\end{figure}

@@

After that, we can generate a table to examine all values of ``descendant" counts in the dataset, along with the number of individuals who have each of those values of ``descendant" counts. Of the 8165 individuals in this dataset, 6252 of them have zero ``descendants", 322 of them have one ``descendant", and 145 of them have two ``descendants". There are only 17 individuals who have more than 30 ``descendants", and there is one individual who has the largest value of 159 ``descendants". We determine that this individual is the prominent British statistician Sir David Cox, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.

@@

We can now visualize how these 159 ``descendants" are related to Sir David Cox by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}. As such, we create Figure \ref{fig:dCox} using the code below.

@@

\begin{figure}%[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{dCox.png}}
    \caption{The 159 academic statistician ``descendants" of Sir David Cox.}
    \label{fig:dCox}
\end{figure}

We see from Figure \ref{fig:dCox} that Sir David Cox had 42 ``children", many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn. Of his ``children", the one who produced the most ``children" of their own was Peter Bloomfield, who has 26 ``children" and 49 ``descendants". In total, Sir David Cox had five generations of academic statistics mentees in this dataset.

@@

At this point, it would be insightful to examine a more detailed view of one of the longest strings of ``parent-child" relationships between Sir David Cox and one of the two individuals who are his fifth generation ``descendants". We do so with the code below, choosing his fifth generation ``descendant" to be Petra Buzkova. We set the \code{fontFace} variable of the \code{plotPath()} to a value of 4, indicating we wish to boldface and italicize the two varieties of interest.

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.7\textwidth]{pathCB}}
    \caption{The shortest path between Sir David Cox and one of his fifth generation ``descendants", Petra Buzkova.}
    \label{fig:pathCB}
\end{figure}

This code results in Figure \ref{fig:pathCB}. We see that the shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional ``parent-child" relationships that span about 55 years. We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than 30 years in degree acquisition, whereas the three statisticians who earned their degrees later in this path only span less than ten years in degree acquisition.

We also notice in Figure \ref{fig:pathCB} that Sir David Cox received his statistics degree in about 1950, and Petra Buzkova received her statistics degree in about 2005. This genealogy only contains historical information about obtained degrees, and does not project into the future. Hence, we can be assured that Petra Buzkova is one of the younger individuals in the dataset, at least in the sense that the youngest individual could only have received his or her degree ten years after Petra Buzkova. However, we cannot be assured that Sir David Cox is one of the oldest individuals in the dataset. As such, it would be informative to superimpose this path of interest onto the entire dataset, using the \code{plotPathOnAll()} function of the \pkg{ggenealogy} package, as we did for the soybean genealogy in Figures \ref{fig:plotTNBin3} and \ref{fig:plotTNBin6}.

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotCBText}}
    \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.}
    \label{fig:plotCBText}
\end{figure}

We see from the resulting Figure \ref{fig:plotCBText} that almost all text labels for individuals who received their graduate-level statistics degrees between 1950 and 2015 are undecipherable. We also see that the year Sir David Cox acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for acquisition of statistics degrees in this dataset occur around 1860. However, the number of individuals who are documented as receiving their statistics degrees between 1860 and 1950 are few enough so that their text labels are somewhat readable.

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotCBNoText}}
    \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200. Individuals on the shortest path are labeled in large and black text and connected by dark green edges; all other individuals are labeled in small and gray text and connected by light gray edges.}
    \label{fig:plotCBNoText}
\end{figure}

In Figure \ref{fig:plotCBNoText}, we can now see each individual on the path of interest, and how their values for the variable year are overlaid on the entire genealogy structure. We can also more clearly see that, even though only ten years span between the youngest individual in the genealogy and Petra Buzkova, there are many individuals in that last decade. Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of acquisition of statistics degrees.

>>>>

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
    \caption{Example pedigree chart from the \pkg{kinship2} package, where the vertical axis denotes generation count. We superimposed green-highlighted individual 215 for explanatory purposes. As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual. Hence, it should be displayed twice on the vertical axis, once for each of its generation counts. However, standard pedigree tools only allow for an individual to be displayed once, leading to ambiguous portrayal of the genealogical dataset.}
    \label{fig:kinshipFig}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{pathTNZB}}
    \caption{Left: The shortest path between varieties \code{Tokyo} and \code{Narow} is strictly composed of a unidirectional sequence of parent-child relationships. Right: The shortest path between varieties \code{Zane} and \code{Bedford} is not strictly composed of unidirectional parent-child relationships; they instead have a cousin-like relationship.}
    \label{fig:pathTNZB}
\end{figure}

@@

\begin{figure}%[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotTNBin3}}
    \caption{The shortest path between \code{Tokyo} and \code{Narow}, superimposed over the data structure, using a bin size of 3.}
    \label{fig:plotTNBin3}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotTNBin6}}
    \caption{The shortest path between \code{Tokyo} and \code{Narow}, superimposed over the data structure, using a bin size of 6.}
    \label{fig:plotTNBin6}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{LeeAD}}
    \caption{Top: All ancestors and descendants of the variety \code{Lee} are shown in this \pkg{ggenealogy} plot. Bottom: We now attempt to mimic the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated. The parenthetical numbers above each node represents the set of generation counts that node is away from the center node \code{Lee}. The presence of red parentheses indicate that the plot on the bottom ambiguously display the example soybean genealogy in the way that the \pkg{ggenealogy} plot on the top can accomplish.}
    \label{fig:Lee}
\end{figure}

@@

After that, we can generate a table to examine all values of ``descendant" counts in the dataset, along with the number of individuals who have each of those values of ``descendant" counts. Of the 8165 individuals in this dataset, 6252 of them have zero ``descendants", 322 of them have one ``descendant", and 145 of them have two ``descendants". There are only 17 individuals who have more than 30 ``descendants", and there is one individual who has the largest value of 159 ``descendants". We determine that this individual is the prominent British statistician Sir \code{David Cox}, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.

@@

We can now visualize how these 159 ``descendants" are related to Sir \code{David Cox} by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}. As such, we create Figure \ref{fig:dCox} using the code below.

@@

\begin{figure}%[h]
    \centering
    \fbox{\includegraphics[width=\textwidth]{dCox.png}}
    \caption{The 159 academic statistician ``descendants" of Sir \code{David Cox}.}
    \label{fig:dCox}
\end{figure}

We see from Figure \ref{fig:dCox} that Sir \code{David Cox} had 42 ``children", many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn. Of his ``children", the one who produced the most ``children" of their own was Peter Bloomfield, who has 26 ``children" and 49 ``descendants". In total, Sir \code{David Cox} had five generations of academic statistics mentees in this dataset.

@@

At this point, it would be insightful to examine a more detailed view of one of the longest strings of ``parent-child" relationships between Sir \code{David Cox} and one of the two individuals who are his fifth generation ``descendants". We do so with the code below, choosing his fifth generation ``descendant" to be \code{Petra Buzkova}. We set the \code{fontFace} variable of the \code{plotPath()} to a value of 4, indicating we wish to boldface and italicize the two varieties of interest.

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.7\textwidth]{pathCB}}
    \caption{The shortest path between Sir \code{David Cox} and one of his fifth generation ``descendants", \code{Petra Buzkova}.}
    \label{fig:pathCB}
\end{figure}

This code results in Figure \ref{fig:pathCB}. We see that the shortest path between Sir \code{David Cox} and \code{Petra Buzkova} is strictly composed of five unidirectional ``parent-child" relationships that span about 55 years. We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than 30 years in degree acquisition, whereas the three statisticians who earned their degrees later in this path only span less than ten years in degree acquisition.

We also notice in Figure \ref{fig:pathCB} that Sir \code{David Cox} received his statistics degree in about 1950, and \code{Petra Buzkova} received her statistics degree in about 2005. This genealogy only contains historical information about obtained degrees, and does not project into the future. Hence, we can be assured that \code{Petra Buzkova} is one of the younger individuals in the dataset, at least in the sense that the youngest individual could only have received his or her degree ten years after \code{Petra Buzkova}. However, we cannot be assured that Sir \code{David Cox} is one of the oldest individuals in the dataset. As such, it would be informative to superimpose this path of interest onto the entire dataset, using the \code{plotPathOnAll()} function of the \pkg{ggenealogy} package, as we did for the soybean genealogy in Figures \ref{fig:plotTNBin3} and \ref{fig:plotTNBin6}.

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotCBText}}
    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.}
    \label{fig:plotCBText}
\end{figure}

We see from the resulting Figure \ref{fig:plotCBText} that almost all text labels for individuals who received their graduate-level statistics degrees between 1950 and 2015 are undecipherable. We also see that the year Sir \code{David Cox} acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for acquisition of statistics degrees in this dataset occur around 1860. However, the number of individuals who are documented as receiving their statistics degrees between 1860 and 1950 are few enough so that their text labels are somewhat readable.

@@

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=\textwidth]{plotCBNoText}}
    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200. Individuals on the shortest path are labeled in large and black text and connected by dark green edges; all other individuals are labeled in small and gray text and connected by light gray edges.}
    \label{fig:plotCBNoText}
\end{figure}

In Figure \ref{fig:plotCBNoText}, we can now see each individual on the path of interest, and how their values for the variable year are overlaid on the entire genealogy structure. We can also more clearly see that, even though only ten years span between the youngest individual in the genealogy and \code{Petra Buzkova}, there are many individuals in that last decade. Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of acquisition of statistics degrees.
